---
title: "Chapter 1"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r}
library(ggplot2)
```

# Worked Problems

## Problem 1.

a) 
```{r}
# The number of phenotypic classes obtained (in general) is given by:
num_classes<-function(n){(2*n)+1}
n<-20 # The number of loci
num_classes(n)
```
b) The probability of finding an individual among the offspring that is heterozygous at all 20 loci is the product of the probabilities at being heterozygous at 1 locus. N.B. We assume that they are unlinked (independent), and the parent is a heterozygote that is reproducing via selfing. Thus for a single locus the frequency of heterozygotes is 2/4 (0.5). So the frequency of heterozygotes at 20 loci is:
```{r}
0.5^n
```

c) The probability of obtaining descendants with a phenotypic value of 10 can be obtained using the binomial theorem.
```{r}
n<-20
k<-10
p<-0.5
q<-1-p
# There are 2n alleles, we choose 10 of them and we want to kno how many have k = 10. This can be obtained with:
choose((2*n),k)*(p^k)*(q^((2*n)-k))
# R also has a function for this
dbinom(x = k, prob = p,size=(2*n))
```

To get the full probability distribution we can compute the probabilities for all k = {0..40}. Again, we are mating two heterozygotes, so the allele frequencies are p = q = 0.5
```{r}
genotypic_values<-seq(0,40)
probabilities<-dbinom(x=genotypic_values,prob=0.5,size=(2*n))
```
Plot this data
```{r}
ggplot()+
  geom_bar(aes(x=genotypic_values,y=probabilities),stat="identity")+
  geom_text(aes(x=genotypic_values,y=probabilities,label=as.character(round(probabilities,5))),hjust=-0.1,vjust=0.5,angle=90)+
  scale_y_continuous(limits=c(0,0.15))+
  xlab("Genotypic Value")+
  ylab("Probability")
```



## Problem 2.
This is the data.
```{r}
bris<-c(24,24,21,21,21,20,19,19,18,24,
  23,25,22,24,22,22,18,20,22,28,
  23,27,23,20,23,21,26,27,27,28,
  26,26,24,26,27,26,20,22,21,18,
  18,21,17,21,18,20,23,22,34,22)
```

The phenotypic variance can be obtained simply from the phenotypic data. And we assume Vg = 4
```{r}
Vp<-var(bris)
Vg<-4
Ve<-Vp-Vg
```

Coefficients of variation for both P and G can be obtained with the mean  of the phenotypic data. Since we assume that the mean of the genotypic values is the same as the mean of the phenotypic values.
```{r}
CVp<-sqrt(Vp)/(mean(bris))
CVg<-sqrt(Vg)/(mean(bris))
cat("Ve = ",Ve, ", CVp = ",CVp, ", CVg = ",CVg,"\n")
```

Self Assessment Questions:
1. False
2. True
3. False
4. False
5. True
6. True
7. False
8. False
9. True
10. True
